```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Week 2: Course Project 1

**Cleaning workspace**
```{r cleaning workspace}
rm(list=ls())
```

**Reading data and showing first 6 rows and number of total rows**
```{r reading data}
activity_data <- read.csv("activity.csv")
head(activity_data)
nrow(activity_data)
```

**Eliminating NAs**
```{r}
activity_data_naomit <- na.omit(activity_data)
```

**Calculating steps per day**
```{r}
library(data.table)
setDT(activity_data_naomit)
head(steps_by_day <- activity_data_naomit[, .(steps_date = sum(steps)), by = date])
```

**Steps histogram**
```{r message=FALSE}
library(ggplot2)
qplot(steps_by_day$steps_date, bins = 10, colour = I("white"))
```

**Mean and median (steps)**
```{r}
mean(steps_by_day$steps_date)
median(steps_by_day$steps_date)
```

**Time series plot, interval vs mean(steps)**
```{r}
(mean_by_interval <- activity_data_naomit[, .(mean_steps=mean(steps)), by = interval])
plot(mean_by_interval, type = "l", xaxt = "n")
axis(1, seq(0, 2355, length.out = 288), labels = mean_by_interval$interval,
     las = 2, cex.axis=.6)
```

**Maximum number of steps**
```{r message=FALSE}
library(dplyr)
arrange(mean_by_interval, desc(mean_steps))[1,]
```

**Total of NAs by steps, date & interval**
```{r}
apply(activity_data, 2, function(M) sum(is.na(M)))
```

**Filling NAs with the mean of the interval**
```{r}
setDT(activity_data)
mean_steps_nas <- merge(activity_data[is.na(steps),"interval"],
                        mean_by_interval, sort = F)$mean_steps
activity_data_full <- copy(activity_data)
activity_data_full[,steps := as.numeric(steps)]
activity_data_full[is.na(steps), steps := mean_steps_nas][]
```

**Total of NAs**
```{r}
apply(activity_data_full, 2, function(M) sum(is.na(M)))
```

**Plotting interval vs mean(steps) with imputed data**
```{r}
mean_by_interval_full <- activity_data_full[, .(mean_steps=mean(steps)),
                                            by = interval]
plot(mean_by_interval_full, type = "l", xaxt = "n")
axis(1, seq(0,2355,length.out = 288),labels = mean_by_interval_full$interval,
     las = 2, cex.axis=.6)
```

**Calculating steps per day**
```{r}
steps_by_day_full <- activity_data_full[, .(steps_date = sum(steps)), by = date]
```

**Histogram of the total numbers of steps**
```{r}
qplot(steps_by_day_full$steps_date, bins = 10, colour = I("white"))
```

**Calculating mean and median of the full data set**
```{r}
mean(steps_by_day_full$steps_date)
median(steps_by_day_full$steps_date)
```

#### The imputed data dosen't impact the mean, nevertheless change the median and the hisogram because there are more values to show.

**Creating group_days factor variable (the names of the week days are in Spanish)**
```{r}
weekend_log <- weekdays(as.Date(activity_data_full$date)) %in%
      sort(weekdays(Sys.Date()+1:7))[c(1,6)]
group_days <- factor(levels = c("weekend", "weekday"))
group_days[weekend_log] <- "weekend"
group_days[!weekend_log] <- "weekday"
table(group_days)
table(weekdays(as.Date(activity_data_full$date)))
activity_data_full[, group_days := group_days][]
```

**Plotting by interval and group_days with lattice**
```{r}
mean_by_interval_full2 <- activity_data_full[, .(mean_steps=mean(steps)),
                                            by = .(interval, group_days)]
library(lattice)
xyplot(mean_steps ~ interval | group_days, data = mean_by_interval_full2,
       layout = c(1, 2), type = "l")
```
